---
title: "Zooplankton in lakes"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

# Markdown!! a way to make live documents

R Notebooks can be a way to do a lot of live documents includin presentaitions, word documents, pdf files, and even web pages.

## Making tables

Today we will use it to make some simple tables

```{r libraries, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
# # Libraries used each time ---------
# uesed each time
library(tidyverse)
library(scales)
library(readxl)
library(janitor)
library(patchwork)
library(plotly)
library(skimr)
library(kableExtra)
```

```{r read_file, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
## read in the file -----
z.df <- read_csv("data/zoops_long.csv") %>% clean_names() %>% 
  mutate(date = as_date(date))
# head(z.df)
```

## Now to make a table of the data we have

all from: <https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html>

```{r include=FALSE}
z_summary.df <- z.df %>% 
  mutate(month = month(date)) %>% 
  group_by(site, year, month, species) %>% 
  summarize(mean_count = mean(number, na.rm = TRUE))
  
```

This is the long summary table

```{r echo=FALSE}
knitr::kable( z_summary.df) %>%  
  kable_classic_2(full_width = FALSE, position = "left", html_font ="Cambria") %>%
  kable_styling(bootstrap_options = c("striped"))
```

So that sort of sucks and not the best - how can we reformat

```{r include=FALSE}
z_summary_wide.df <- z_summary.df %>% 
  pivot_wider(
    names_from = species,
    values_from = mean_count
  )
```

So here is the wide table

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
knitr::kable( z_summary_wide.df) %>%
  kable_classic_2(full_width = FALSE, position = "left", html_font ="Cambria") %>%
  kable_styling(bootstrap_options = c("striped"))
```
